Relocating a Cluster of Earthquakes Using a Single Seismic Station

### 2013-01-02 paper2/paper2.tex
a number of modifications after firs feedback from BSSA editorial office

%\documentclass[extra,mreferee]{gji}
\documentclass[extra, onecolumn, doublespacing]{gji}
%\documentclass[extra]{gji}

\bibliographystyle{gji}
\usepackage{timet}
\usepackage[dvips]{graphicx}

\title{Relocating a Cluster of Earthquakes Using a Single Seismic Station}
\author[D.J. Robinson, M. Sambridge, R. Snieder and J Hauser]
  {David J. Robinson$^{1,2}$, Malcolm Sambridge$^{1}$, Roel Snieder$^{3}$, Juerg Hauser$^{4}$ \\
  $^1$ Research School of Earth Sciences, Australian National
    University, Canberra ACT \emph{0200}, Australia, E-mail: david.robinson@anu.edu.au \\
$^2$  Risk Analysis Methods, Geoscience Australia, GPO Box \emph{383} Canberra ACT \emph{2601} Australia, E-mail: david.robinson@ga.gov.au\\
$^3$ Center for Wave Phenomena and Department of Geophysics, Colorado School of Mines, Golden CO \emph{80401-1887}, USA \\
$^4$ CSIRO Earth Science and Resource Engineering, Dick Perry Av, Kensington WA \emph{6151} Australia}

\date{Received 2012 Month Day; in original form 2012 Month Day}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\volume{200}
\pubyear{2012}

@@

\label{firstpage}

\maketitle

\begin{summary}
Coda waves arise from scattering to form the later arriving
components of a seismogram. Coda wave interferometry is an emerging
tool for constraining earthquake source properties from the
intereference pattern of coda waves between nearby events. A new
earthquake location algorithm is derived which relies on coda wave
based probabilistic estimates of earthquake separation. The
algorithm can be used with coda waves alone or in tandem with travel
time data. Synthetic examples in 2D and 3D and real earthquakes on
the Calaveras Fault, California are used to demonstrate the
potential of coda waves for locating poorly recorded earthquakes. It
is demonstrated that coda wave interferometry: (a) outperforms
traditional earthquake location techniques when the number of
stations is small; (b) is self-consistent across a broad range of
station situations; and (c) can be used with a single station to
locate earthquakes.

\end{summary}

@@

\section{Introduction}

@@

Absolute location describes the location of an earthquake with
respect to a global reference such as latitude, longitude (or
easting/northing) and depth. Uncertainties associated with absolute
locations are influenced by source to station distances, the number
of stations and their geometry, signal-to-noise, clarity of onsets
and accuracy of the velocity model used in computing travel times.
Uncertainties in absolute location are typically of the order of
several kilometers because they are susceptible to uncertainty in
the velocity structure along the entire path between the source and
receiver. For example, \citet{dr_Shearer99a} states that location
uncertainty in the ISC (International Sesimological
Centre\footnote{http://www.isc.ac.uk}) and PDE (National Earthquake
Information
Center\footnote{http://earthquake.usgs.gov/earthquakes/eqarchives/epic/})
catalogues are generally around 25\,km horizontally and at least
25\,km in depth.\footnote{Here the depth uncertainties of 25\,km
assume the use of depth dependent phases such as $pP$. Without such
phases the uncertainty is higher.} \citet{dr_Bondar04a} demonstrate
that at the local scale, absolute locations are accurate to within
5\,km with a 95\% confidence level when local networks meet a number
of station related criteria. Such errors are too large for many
applications, particularly those focussed on imaging rupture
surfaces from aftershock sequences.

Relative earthquake location involves locating a group of
earthquakes with respect to one another and was first introduced by
\citet{dr_Douglas67a} who developed the technique commonly known as
joint hypocenter determination.\footnote{\citet{dr_Douglas67a}
originally used the term joint epicentre determination. However, he
was solving for hypocentre.} In principle, relative locations can be
computed by differencing absolute locations. However,
\citet{dr_Pavlis92a} shows that inadequate knowledge of velocity
structure leads to systematic biases when relative positions are
computed in this way. To reduce errors from unknown velocity
structure, relative location techniques compute locations directly
from travel time differences between two waveforms
\citep{dr_Ito85a, dr_Got94a, dr_Nadeau97a, dr_Waldhauser99a}. By doing
so, they remove errors associated with velocity variations outside
the local region, because such variations influence all waveforms in
the same manner \citep{dr_Shearer99a}.

@@

Coda refers to  later arriving waves in the seismogram that arise
from scattering  \citep{dr_Aki69a,dr_Snieder99a,dr_Snieder06a}. Coda
waves are ignored in most seismological applications due to the
complexity involved in constraining complex hetergeneous velocity models
in real settings. In this paper we
develop an approach for locating earthquakes using coda waves.
\citet{dr_Snieder05a} demonstrate that the coda of two earthquakes
can be used to estimate the separation between them. Their
technique, known as coda wave interferometry (CWI), is based on the
interference pattern between the coda waves. Unlike travel time
based location techniques, CWI
 does not require multiple stations or good azimuthal coverage.
In fact, it is possible to obtain estimates of separation using a single
station \citep{dr_Robinson07b}. This makes CWI particularly interesting
for regions where station density is low such as intraplate settings. In this
paper we demonstrate how CWI separation estimates can be used to constrain
location with data from a single station.
Our technique can be used on coda waves alone or in combination with
travel times. We begin by introducing the theory of CWI based earthquake location
(Section \ref{sec:theory}). This is followed by a demonstration of capability using
synthetic examples (Section \ref{sec:benchmarking}) and application to earthquakes
on the Calaveras fault, California using CWI alone (Section \ref{sec:CalaverasLoc-CWIonly}
and CWI in combination with travel time constraints (Section \ref{sec:CalaverasLoc-CWIandTT}).

\section{Theory}
\label{sec:theory}

\citet{dr_Snieder05a} introduce a CWI based estimator of source
separation $\delta_{CWI}$ between two earthquakes
\begin{equation}
\label{eq:Theory-dcwi_g_sigmat}
\delta_{CWI}^2 = g(\alpha,\beta)\sigma_\tau^2,
\end{equation}
where $\sigma_\tau$ is the standard deviation of the travel time
perturbation between the coda waves of two earthquakes, and $\alpha$
and $\beta$ are the near-source $P$ and $S$ wave velocities,
respectively. The function $g$ depends on the type of excitation
(explosion, point force, double couple) and on the direction of
source displacement relative to the point force or double couple.
For example, for two double couple sources displaced in the fault
plane,
\begin{equation}
\label{eq:LitReview-g-3d-dc}
g(\alpha,\beta) = 7
\frac{\left(\frac{2}{\alpha^6}+\frac{3}{\beta^6}\right)}{\left(\frac{6}{\alpha^8}+\frac{7}{\beta^8}\right)},
\end{equation}
whereas, for two point sources in a 2D acoustic medium
\begin{equation}
\label{eq:LitReview-g-2d-ps} g(\alpha,\beta) = 2 \alpha^2
\end{equation}
\citep{dr_Snieder05a}. In this paper we use equation
\ref{eq:LitReview-g-3d-dc} which assumes that the source mechanism
of both events are identical, an assumption likely to be true for
events in the same fault plane. \cite{dr_Robinson07c} explore the
impact of a change in mechanism.

The $\sigma_\tau$ in equation (\ref{eq:Theory-dcwi_g_sigmat}) is
related to the maximum of the cross correlation between the coda of
the two waveforms, $R_{max}$, and hence can be computed directly
from the recorded data. The original formulation by
\citet{dr_Snieder05a} used a second-order Taylor series expansion of
the waveform autocorrelation function to relate $\sigma_\tau$ and
$R_{max}$ by
\begin{equation}
\label{eq:Rmax-disp-orig} R_{max}^{(t,t_w)} = 1 - \frac{1}{2}
\overline{\omega^2} \sigma_\tau^2,
\end{equation}
In this paper we use the autocorrelation approach of
\citet{dr_Robinson11a} to relate the parameters directly and we
apply a restricted time lag search when evaluating $R_{max}$. These
extensions to the original technique of \citet{dr_Snieder05a}
increase the range of applicability of CWI by 50\% (i.e. from 300 to
450\,m separation for 1 to 5\,Hz filtered coda waves).

\citet{dr_Robinson11a} show that CWI leads to
probabilistic constraints on source separation and introduce a
Bayesian approach for describing the probability of true separation
given the CWI data. Their approach is summarised by
\begin{equation}
\label{eq:-bayesian-noisyform}
P(\widetilde{\delta}_t|\widetilde{\delta}_{CWIN}) \propto P(\widetilde{\delta}_{CWIN}|\widetilde{\delta}_t)
\times P(\widetilde{\delta}_t)
\end{equation}
where  $P(\widetilde{\delta}_t|\widetilde{\delta}_{CWIN})$ is the
posterior function indicating the probability of true separation
$\widetilde{\delta}_t$ given the noisy CWI separation estimates
$\widetilde{\delta}_{CWIN}$;
$P(\widetilde{\delta}_{CWIN}|\widetilde{\delta}_t)$ is the
likelihood function (or forward model) giving the probability that
the separation estimates $\widetilde{\delta}_{CWIN}$ would be
observed if the true separation was $\widetilde{\delta}_t$; and
$P(\widetilde{\delta}_t)$ is the prior PDF accounting for all
a-priori information. The tilde above the separation parameters in
equation (\ref{eq:-bayesian-noisyform}) indicates the use of a
wavelength normalised separation parameter
\begin{equation}
\label{eq-normalisation-eqn}
\widetilde{\delta} = \frac{\delta}{\lambda_{d}},
\end{equation}
which measures separation ($\delta$ = $\delta_{CWIN}$ or $\delta_t$)
with respect to dominant wavelength $\lambda_{d}$. In this paper we
consider a uniform prior over appropriate bounds to ensure that the
posterior function is dominated by the recorded data.
The procedure for computing the likelihood
$P(\widetilde{\delta}_{CWIN}|\widetilde{\delta}_t)$ is derived by
\citet{dr_Robinson11a} and summarised in Appendix
\ref{sec-Appendix-noisylikelihood}. With these two pieces in place
we can compute the posterior
$P(\widetilde{\delta}_t|\widetilde{\delta}_{CWIN})$ (or PDF) for the
separation between any pair of events directly from their coda
waves.

We seek a probability density function (PDF) which links
 individual pairwise posteriors $P(\widetilde{\delta}_t|\widetilde{\delta}_{CWIN})$
to describe the location of multiple events whose maximum corresponds to
 the most probable combination of locations.
More importantly however, the PDF shall
quantify location uncertainty and provide information
on the degree to which individual events are constrained
 by the data. For convenience, we begin with three earthquakes having locations
$\mathbf{e}_1$, $\mathbf{e}_2$ and $\mathbf{e}_3$. Using a Bayesian
formulation we write
\begin{equation}
\label{eq-Bayes-three-locations}
P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3| \mathbf{d}) \propto P( \mathbf{d}|\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3)
\times P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3),
\end{equation}
where $P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3 | \mathbf{d})$,
$P(\mathbf{d}|\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3)$ and
$P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3)$ are the posterior,
likelihood and prior functions, respectively. In equation
(\ref{eq-Bayes-three-locations}) $\mathbf{d}$ represents
observations that constrain the locations. They can be any
combination of travel times, geodetic information  or CWI
separations. For example, if coda waves are used we have
$P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3 |
\widetilde{\mitbf{\delta}}_{CWIN})$ and
$P(\widetilde{\mitbf{\delta}}_{CWIN}|\mathbf{e}_1, \mathbf{e}_2,
\mathbf{e}_3)$, where $\widetilde{\mitbf{\delta}}_{CWIN}$ are the
wavelength normalised separation estimates. Alternatively, if we use
CWI and travel time data we may write $P(\mathbf{e}_1, \mathbf{e}_2,
\mathbf{e}_3 | \widetilde{\mitbf{\delta}}_{CWIN},
\mathbf{\Delta}_{TT})$ and
$P(\widetilde{\mitbf{\delta}}_{CWIN},\mathbf{\Delta}_{TT}|\mathbf{e}_1,
\mathbf{e}_2, \mathbf{e}_3)$ where $\mathbf{\Delta}_{TT}$ represent
travel time differences. In the following derivation and in Sections
\ref{sec:benchmarking} and \ref{sec:CalaverasLoc-CWIonly} we focus
on the constraints imposed by coda waves, whereas in Section
\ref{sec:CalaverasLoc-CWIandTT} we demonstrate how CWI and travel
time data can be combined.

For three earthquakes we have likelihoods;
$P(\widetilde{\delta}_{CWIN,12}|\mathbf{e}_1, \mathbf{e}_2)$,
$P(\widetilde{\delta}_{CWIN,13}|\mathbf{e}_1, \mathbf{e}_3)$
and
$P(\widetilde{\delta}_{CWIN,23}|\mathbf{e}_2, \mathbf{e}_3)$.
In writing
these likelihoods we have replaced the conditional term on separation
$\widetilde{\delta}_t$ with the locations (e.g. $\mathbf{e}_1$ and $\mathbf{e}_2$).
This can be done because
knowledge of location translates to separation.\footnote{Note that the reverse is
not true. That is, knowledge of separation between a single event pair
does not uniquely translate to location but rather places a non-unique constraint on
location.}
Furthermore, since the pairwise functions are independent the joint
likelihood becomes
\begin{equation}
\begin{array}{l}
\label{eq:-joint-likelihood4three}
P(\widetilde{\mathbf{\delta}}_{CWIN} | \mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3) =
P(\widetilde{\delta}_{CWIN,12} | \mathbf{e}_1, \mathbf{e}_2) \\
\hspace{2em} \times P(\widetilde{\delta}_{CWIN,13} | \mathbf{e}_1, \mathbf{e}_3)
\times  P(\widetilde{\delta}_{CWIN,23} | \mathbf{e}_2, \mathbf{e}_3).
\end{array}
\end{equation}
Similarly, the earthquake locations are independent and the
joint prior becomes
\begin{equation}
\label{eq:-joint-prior4three} P(\mathbf{e}_1, \mathbf{e}_2,
\mathbf{e}_3) = P(\mathbf{e}_1) \times P(\mathbf{e}_2) \times
P(\mathbf{e}_3).
\end{equation}
Combining equations (\ref{eq:-joint-likelihood4three}) and
(\ref{eq:-joint-prior4three}) gives the joint posterior function
\begin{equation}
\label{eq-3eqposterior}
\begin{array}{l}
P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3 | \widetilde{\mitbf{\delta}}_{CWIN}) = c \displaystyle \prod_{i=1}^3 P(\mathbf{e}_i) \\
\hspace{4em}  \times \displaystyle \prod_{i=1}^2 \prod_{j=i+1}^3
P(\widetilde{\delta}_{CWIN,ij}|\mathbf{e}_i,\mathbf{e}_j)
\end{array}
\end{equation}
for three events.

A detailed understanding of the location of a single event (e.g. $\mathbf{e}_2$) is obtained by computing the
marginal
\begin{equation}
\label{eq:-E2-marginal}
P(\mathbf{e}_2|\mathbf{\delta}_{CWIN}) = \int \int P(\mathbf{e}_1, \mathbf{e}_2,
 \mathbf{e}_3 | \widetilde{\mathbf{\delta}}_{CWIN}) d\mathbf{e}_1
 d\mathbf{e}_3,
\end{equation}
where the intergral is taken over all plausible locations for
$\mathbf{e}_1$ and $\mathbf{e}_3$. Alternatively, we can compute the
marginal for a single event coordinate by integrating the posterior
over all events and remaining coordinates for the chosen earthquake.
Evaluation of the normalizing constant $c$ in equation (\ref{eq-3eqposterior}) involves
finding the integral of the posterior function
over all plausible locations. In many applications the constant
of proportionality $c$ can be ignored. For example, it is not
required when seeking the combination of locations which maximise
 the posterior function, nor in Bayesian sampling algorithms such as Markov-chain Monte-Carlo
techniques which only require evaluation of a function proportional to the PDF.

>>>>

%\documentclass[extra,mreferee]{gji}
%\documentclass[extra, onecolumn, doublespacing]{gji}
%\documentclass[extra]{gji}
%\RequirePackage{lineno}
\documentclass[12pt,double]{article}

%\bibliographystyle{agufull08} % Uncomment when generating bibliography
\usepackage{timet}
\usepackage[dvips]{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{rotating}
\usepackage{natbib}
\usepackage{lineno}
%\usepackage{titlesec}
\linenumbers

\linespread{1.5}

\newcommand{\mitbf}[1]{  
  \hbox{\mathversion{bold}$#1$}}  
  \newcommand{\rmn}[1] {{\mathrm #1}}  
  \newcommand{\itl}[1] {{\mathit #1}}  
  \newcommand{\bld}[1] {{\mathbf #1}}

%\titleformat{\section}
%{\bf}{\thesection}{1em}{\MakeUppercase{#1}}

%\date{Received 2012 Month Day; in original form 2012 Month Day}
%\pagerange{\pageref{firstpage}--\pageref{lastpage}}
%\volume{200}
%\pubyear{2012}

@@

\begin{centering}
\LARGE{\textbf{Relocating a Cluster of Earthquakes Using a Single Seismic Station}} \\  
\end{centering}
\vspace{3em}
\begin{centering}
David J. Robinson, Malcolm Sambridge, Roel Snieder, Juerg Hauser \\
\end{centering}

\vspace{7em}
\begin{tabular}{lp{0.6\textwidth}}
 & \\
David Robinson: & Earth Monitoring and Hazard Group \\
 &  Geoscience Australia \\
 & GPO Box \emph{383} \\
 & Canberra ACT \emph{2601} Australia \\
 & E-mail: david.robinson@ga.gov.au\\
\end{tabular}

\clearpage
\label{firstpage}

%\maketitle

\vspace{2em}
\begin{centering}
\section*{Abstract}
\end{centering}
Coda waves arise from scattering to form the later arriving
components of a seismogram. Coda wave interferometry is an emerging
tool for constraining earthquake source properties from the
intereference pattern of coda waves between nearby events. A new
earthquake location algorithm is derived which relies on coda wave
based probabilistic estimates of earthquake separation. The
algorithm can be used with coda waves alone or in tandem with travel
time data. Synthetic examples in 2D and 3D and real earthquakes on
the Calaveras Fault, California are used to demonstrate the
potential of coda waves for locating poorly recorded earthquakes. It
is demonstrated that coda wave interferometry: (a) outperforms
traditional earthquake location techniques when the number of
stations is small; (b) is self-consistent across a broad range of
station situations; and (c) can be used with a single station to
locate earthquakes.

@@

\vspace{2em}
\begin{centering}
\section*{Introduction} 
\end{centering}

@@

Absolute location describes the location of an earthquake with
respect to a global reference such as latitude, longitude (or
easting/northing) and depth. Uncertainties associated with absolute
locations are influenced by source to station distances, the number
of stations and their geometry, signal-to-noise, clarity of onsets
and accuracy of the velocity model used in computing travel times.
Uncertainties in absolute location are typically of the order of
several kilometers because they are susceptible to uncertainty in
the velocity structure along the entire path between the source and
receiver. For example, \citet{dr_Shearer99a} states that location
uncertainty in the ISC (International Sesimological
Centre) and PDE (National Earthquake
Information
Center)
catalogues are generally around 25\,km horizontally and at least
25\,km in depth (Here the depth uncertainties of 25\,km
assume the use of depth dependent phases such as $pP$. Without such
phases the uncertainty is higher). \citet{dr_Bondar04a} demonstrate
that at the local scale, absolute locations are accurate to within
5\,km with a 95\% confidence level when local networks meet a number
of station related criteria. Such errors are too large for many
applications, particularly those focussed on imaging rupture
surfaces from aftershock sequences.

Relative earthquake location involves locating a group of
earthquakes with respect to one another and was first introduced by
\citet{dr_Douglas67a} who developed the technique commonly known as
joint hypocenter determination (\citet{dr_Douglas67a}
originally used the term joint epicentre determination. However, he
was solving for hypocentre). In principle, relative locations can be
computed by differencing absolute locations. However,
\citet{dr_Pavlis92a} shows that inadequate knowledge of velocity
structure leads to systematic biases when relative positions are
computed in this way. To reduce errors from unknown velocity
structure, relative location techniques compute locations directly
from travel time differences between two waveforms
\citep{dr_Ito85a, dr_Got94a, dr_Nadeau97a, dr_Waldhauser99a}. By doing
so, they remove errors associated with velocity variations outside
the local region, because such variations influence all waveforms in
the same manner \citep{dr_Shearer99a}.

@@

Coda refers to  later arriving waves in the seismogram that arise
from scattering  \citep{dr_Aki69a,dr_Snieder99a,dr_Snieder06a}. Coda
waves are ignored in most seismological applications due to the
complexity involved in constraining complex hetergeneous velocity models
in real settings. In this paper we
develop an approach for locating earthquakes using coda waves.
\citet{dr_Snieder05a} demonstrate that the coda of two earthquakes
can be used to estimate the separation between them. Their
technique, known as coda wave interferometry (CWI), is based on the
interference pattern between the coda waves. Unlike travel time
based location techniques, CWI
 does not require multiple stations or good azimuthal coverage.
In fact, it is possible to obtain estimates of separation using a single
station \citep{dr_Robinson07b}. This makes CWI particularly interesting
for regions where station density is low such as intraplate settings. In this
paper we demonstrate how CWI separation estimates can be used to constrain
location with data from a single station.
Our technique can be used on coda waves alone or in combination with
travel times. We begin by introducing the theory of CWI based earthquake location. 
This is followed by a demonstration of capability using
synthetic examples and application to earthquakes
on the Calaveras fault, California using CWI alone
and CWI in combination with travel time constraints.

\vspace{2em}
\begin{centering}
\section*{Theory}
\end{centering}
\label{sec:theory}

\citet{dr_Snieder05a} introduce a CWI based estimator of source
separation $\delta_{CWI}$ between two earthquakes
\begin{linenomath*} \begin{equation}
\label{eq:Theory-dcwi_g_sigmat}
\delta_{CWI}^2 = g(\alpha,\beta)\sigma_\tau^2,
\end{equation} \end{linenomath*}
where $\sigma_\tau$ is the standard deviation of the travel time
perturbation between the coda waves of two earthquakes, and $\alpha$
and $\beta$ are the near-source $P$ and $S$ wave velocities,
respectively. The function $g$ depends on the type of excitation
(explosion, point force, double couple) and on the direction of
source displacement relative to the point force or double couple.
For example, for two double couple sources displaced in the fault
plane,
\begin{linenomath*} \begin{equation}
\label{eq:LitReview-g-3d-dc}
g(\alpha,\beta) = 7
\frac{\left(\frac{2}{\alpha^6}+\frac{3}{\beta^6}\right)}{\left(\frac{6}{\alpha^8}+\frac{7}{\beta^8}\right)},
\end{equation} \end{linenomath*}
whereas, for two point sources in a 2D acoustic medium
\begin{linenomath*} \begin{equation}
\label{eq:LitReview-g-2d-ps} g(\alpha,\beta) = 2 \alpha^2
\end{equation} \end{linenomath*}
\citep{dr_Snieder05a}. In this paper we use equation
\ref{eq:LitReview-g-3d-dc} which assumes that the source mechanism
of both events are identical, an assumption likely to be true for
events in the same fault plane. \cite{dr_Robinson07c} explore the
impact of a change in mechanism.

The $\sigma_\tau$ in equation (\ref{eq:Theory-dcwi_g_sigmat}) is
related to the maximum of the cross correlation between the coda of
the two waveforms, $R_{max}$, and hence can be computed directly
from the recorded data. The original formulation by
\citet{dr_Snieder05a} used a second-order Taylor series expansion of
the waveform autocorrelation function to relate $\sigma_\tau$ and
$R_{max}$ by
\begin{linenomath*} \begin{equation}
\label{eq:Rmax-disp-orig} R_{max}^{(t,t_w)} = 1 - \frac{1}{2}
\overline{\omega^2} \sigma_\tau^2,
\end{equation} \end{linenomath*}
In this paper we use the autocorrelation approach of
\citet{dr_Robinson11a} to relate the parameters directly and we
apply a restricted time lag search when evaluating $R_{max}$. These
extensions to the original technique of \citet{dr_Snieder05a}
increase the range of applicability of CWI by 50\% (i.e. from 300 to
450\,m separation for 1 to 5\,Hz filtered coda waves).

\citet{dr_Robinson11a} show that CWI leads to
probabilistic constraints on source separation and introduce a
Bayesian approach for describing the probability of true separation
given the CWI data. Their approach is summarised by
\begin{linenomath*} \begin{equation}
\label{eq:-bayesian-noisyform}
P(\widetilde{\delta}_t|\widetilde{\delta}_{CWIN}) \propto P(\widetilde{\delta}_{CWIN}|\widetilde{\delta}_t)
\times P(\widetilde{\delta}_t)
\end{equation} \end{linenomath*}
where  $P(\widetilde{\delta}_t|\widetilde{\delta}_{CWIN})$ is the
posterior function indicating the probability of true separation
$\widetilde{\delta}_t$ given the noisy CWI separation estimates
$\widetilde{\delta}_{CWIN}$;
$P(\widetilde{\delta}_{CWIN}|\widetilde{\delta}_t)$ is the
likelihood function (or forward model) giving the probability that
the separation estimates $\widetilde{\delta}_{CWIN}$ would be
observed if the true separation was $\widetilde{\delta}_t$; and
$P(\widetilde{\delta}_t)$ is the prior PDF accounting for all
a-priori information. The tilde above the separation parameters in
equation (\ref{eq:-bayesian-noisyform}) indicates the use of a
wavelength normalised separation parameter
\begin{linenomath*} \begin{equation} 
\label{eq-normalisation-eqn}
\widetilde{\delta} = \frac{\delta}{\lambda_{d}},
\end{equation} \end{linenomath*}
which measures separation ($\delta$ = $\delta_{CWIN}$ or $\delta_t$)
with respect to dominant wavelength $\lambda_{d}$. In this paper we
consider a uniform prior over appropriate bounds to ensure that the
posterior function is dominated by the recorded data.
The procedure for computing the likelihood
$P(\widetilde{\delta}_{CWIN}|\widetilde{\delta}_t)$ is derived by
\citet{dr_Robinson11a} and summarised in Appendix
\ref{sec-Appendix-noisylikelihood}. With these two pieces in place
we can compute the posterior
$P(\widetilde{\delta}_t|\widetilde{\delta}_{CWIN})$ (or PDF) for the
separation between any pair of events directly from their coda
waves.

We seek a probability density function (PDF) which links
 individual pairwise posteriors $P(\widetilde{\delta}_t|\widetilde{\delta}_{CWIN})$
to describe the location of multiple events whose maximum corresponds to
 the most probable combination of locations.
More importantly however, the PDF shall
quantify location uncertainty and provide information
on the degree to which individual events are constrained
 by the data. For convenience, we begin with three earthquakes having locations
$\mathbf{e}_1$, $\mathbf{e}_2$ and $\mathbf{e}_3$. Using a Bayesian
formulation we write
\begin{linenomath*} \begin{equation}
\label{eq-Bayes-three-locations}
P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3| \mathbf{d}) \propto P( \mathbf{d}|\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3)
\times P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3),
\end{equation} \end{linenomath*}
where $P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3 | \mathbf{d})$,
$P(\mathbf{d}|\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3)$ and
$P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3)$ are the posterior,
likelihood and prior functions, respectively. In equation
(\ref{eq-Bayes-three-locations}) $\mathbf{d}$ represents
observations that constrain the locations. They can be any
combination of travel times, geodetic information  or CWI
separations. For example, if coda waves are used we have
$P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3 |
\widetilde{\mitbf{\delta}}_{CWIN})$ and
$P(\widetilde{\mitbf{\delta}}_{CWIN}|\mathbf{e}_1, \mathbf{e}_2,
\mathbf{e}_3)$, where $\widetilde{\mitbf{\delta}}_{CWIN}$ are the
wavelength normalised separation estimates. Alternatively, if we use
CWI and travel time data we may write $P(\mathbf{e}_1, \mathbf{e}_2,
\mathbf{e}_3 | \widetilde{\mitbf{\delta}}_{CWIN},
\mathbf{\Delta}_{TT})$ and
$P(\widetilde{\mitbf{\delta}}_{CWIN},\mathbf{\Delta}_{TT}|\mathbf{e}_1,
\mathbf{e}_2, \mathbf{e}_3)$ where $\mathbf{\Delta}_{TT}$ represent
travel time differences. In the following derivation and in 
Synthetic Experiements and Relocating Earthquakes on the Calaveras Fault we focus
on the constraints imposed by coda waves, whereas in 
Combining Travel Time and CWI Constraints we demonstrate how CWI and travel
time data can be combined.

For three earthquakes we have likelihoods;
$P(\widetilde{\delta}_{CWIN,12}|\mathbf{e}_1, \mathbf{e}_2)$,
$P(\widetilde{\delta}_{CWIN,13}|\mathbf{e}_1, \mathbf{e}_3)$
and
$P(\widetilde{\delta}_{CWIN,23}|\mathbf{e}_2, \mathbf{e}_3)$.
In writing
these likelihoods we have replaced the conditional term on separation
$\widetilde{\delta}_t$ with the locations (e.g. $\mathbf{e}_1$ and $\mathbf{e}_2$).
This can be done because
knowledge of location translates to separation.Note that the reverse is
not true. That is, knowledge of separation between a single event pair
does not uniquely translate to location but rather places a non-unique constraint on
location.
Furthermore, since the pairwise functions are independent the joint
likelihood becomes
\begin{linenomath*} \begin{equation}
\begin{array}{l}
\label{eq:-joint-likelihood4three}
P(\widetilde{\mathbf{\delta}}_{CWIN} | \mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3) =
P(\widetilde{\delta}_{CWIN,12} | \mathbf{e}_1, \mathbf{e}_2) \\
\hspace{2em} \times P(\widetilde{\delta}_{CWIN,13} | \mathbf{e}_1, \mathbf{e}_3)
\times  P(\widetilde{\delta}_{CWIN,23} | \mathbf{e}_2, \mathbf{e}_3).
\end{array}
\end{equation} \end{linenomath*}
Similarly, the earthquake locations are independent and the
joint prior becomes
\begin{linenomath*} \begin{equation}
\label{eq:-joint-prior4three} P(\mathbf{e}_1, \mathbf{e}_2,
\mathbf{e}_3) = P(\mathbf{e}_1) \times P(\mathbf{e}_2) \times
P(\mathbf{e}_3).
\end{equation} \end{linenomath*}
Combining equations (\ref{eq:-joint-likelihood4three}) and
(\ref{eq:-joint-prior4three}) gives the joint posterior function
\begin{linenomath*} \begin{equation}
\label{eq-3eqposterior}
\begin{array}{l}
P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3 | \widetilde{\mitbf{\delta}}_{CWIN}) = c \displaystyle \prod_{i=1}^3 P(\mathbf{e}_i) \\
\hspace{4em}  \times \displaystyle \prod_{i=1}^2 \prod_{j=i+1}^3
P(\widetilde{\delta}_{CWIN,ij}|\mathbf{e}_i,\mathbf{e}_j)
\end{array}
\end{equation} \end{linenomath*}
for three events.

A detailed understanding of the location of a single event (e.g. $\mathbf{e}_2$) is obtained by computing the
marginal
\begin{linenomath*} \begin{equation}
\label{eq:-E2-marginal}
P(\mathbf{e}_2|\mathbf{\delta}_{CWIN}) = \int \int P(\mathbf{e}_1, \mathbf{e}_2,
 \mathbf{e}_3 | \widetilde{\mathbf{\delta}}_{CWIN}) d\mathbf{e}_1
 d\mathbf{e}_3,
\end{equation} \end{linenomath*}
where the intergral is taken over all plausible locations for
$\mathbf{e}_1$ and $\mathbf{e}_3$. Alternatively, we can compute the
marginal for a single event coordinate by integrating the posterior
over all events and remaining coordinates for the chosen earthquake.
Evaluation of the normalizing constant $c$ in equation (\ref{eq-3eqposterior}) involves
finding the integral of the posterior function
over all plausible locations. In many applications the constant
of proportionality $c$ can be ignored. For example, it is not
required when seeking the combination of locations which maximise
 the posterior function, nor in Bayesian sampling algorithms such as Markov-chain Monte-Carlo
techniques which only require evaluation of a function proportional to the PDF.

Extending to $n$ events we get the posterior function
\begin{linenomath*} \begin{equation}
\begin{array}{l}
\label{posterior-n-events}
P(\mathbf{e}_1,...,\mathbf{e}_n | \widetilde{\mathbf{\delta}}_{CWIN}) = c \displaystyle \prod_{i=1}^n P(\mathbf{e}_i) \\
\hspace{4em} \displaystyle  \times \prod_{i=1}^{n-1} \prod_{j=i+1}^n P(\widetilde{\delta}_{CWIN,ij}|\mathbf{e}_i,\mathbf{e}_j).
\end{array}
\end{equation} \end{linenomath*}
When evaluating equation (\ref{posterior-n-events})
over a range of locations it is necessary to compute and multiply
many numbers close to zero. This is because the PDFs tend to zero as
the locations get less likely (i.e. near the boundaries of the
plausible region). Such calculations are prone to truncation errors
and so we work with the negative logarithm
\begin{linenomath*} \begin{equation}
\label{eq:-negative-log-part1}
L(\mathbf{e}_1, \mathbf{e}_2, ..., \mathbf{e}_n) = -ln\left[ P(\mathbf{e}_1,...,\mathbf{e}_n | \widetilde{\mathbf{\delta}}_{CWIN} )  \right]
\end{equation} \end{linenomath*}
or
\begin{linenomath*} \begin{equation}
\begin{array}{l}
\label{eq:-negative-log}
L(\mathbf{e}_1, \mathbf{e}_2, ..., \mathbf{e}_n) =
-ln\left[ c \right] - \displaystyle \sum_{i=1}^n ln\left[P(\mathbf{e}_i)\right] \\
\hspace{3em} - \displaystyle \sum_{i=1}^{n-1} \sum_{j=i+1}^n ln\left[P(\widetilde{\delta}_{CWIN,ij}|\mathbf{e}_i,\mathbf{e}_j)\right].
\end{array}
\end{equation} \end{linenomath*}
The logarithm improves numerical stability by replacing products
with summations. The negative facilitates the use of optimisation
algorithms that are designed to minimise an objective function.
